\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-21 paper/experiments.tex
paper: added some preliminary experiments

\begin{table*}
  \centerfloat
  \input{runtimes}
    \caption{Base error, percent change in error and and running time for queries
    with different imputation levels. Base error is the root-mean-square error (RMSE) between the query run on clean
    data and the query run on dirty data without imputation. Change in error is relative to the base error.}
  \label{fig:experiments}
\end{table*}

\todobox{actually run experiments and rewrite this section}{
Table~\ref{fig:experiments} provides a summary of our experimental results.
We run each query in four configurations: without imputation on clean data, without imputation on dirty data, and with imputation on dirty data for $\alpha = 0$ and $\alpha = 1$.
When a query returns a relation of the same ``shape'' (same number of tuples and same schema) on both clean and dirty data, we can compute an error value.
For our experiments, we report root-mean-square error (RMSE) for the base error (i.e. the error in the dirty data without imputation with respect to the clean data) and the change
in RMSE for the imputed plans.
For imputations, we run queries using the two extreme values of $\alpha$: 0 and 1, representing a focus on runtime and on information loss, respectively.
For each query, the table displays the base error, the change in error in both imputation scenarios, and the running time in seconds.

Planning times are not shown, as they are fractions of a second for all of the queries that we tested.
The overall low planning times highlight the practicality of our algorithm, despite the exponential complexity.
Running times clearly vary based on $\alpha$, allowing users to tune their query performance as desired.

We can group our queries into five categories: sum/count queries, average queries, min/max queries, selection queries, and join queries on synthetic data.

\begin{itemize}
\item For sum and count queries (\ref{q2}, \ref{q3}, \ref{q5}), imputation does a good job of reducing the error.
For each of these queries, we can reduce total error by several orders of magnitude.
Imputation performs very well in these cases, even for data which is missing uniformly at random, because each missing value directly increases the error of the result.

\item For average queries (\ref{q4}, \ref{q6}, \ref{q7}), imputation does not reduce the error, mainly because running these queries on dirty data incurs only a small penalty.
This is because, in the case of average, removing data uniformly at random does not change the mean.
If we were to remove data in a biased way, imputation would likely perform better because there would be more error for it to correct.

\item For a min query (\ref{q8}), it is unlikely that imputation will ever help, because by definition these queries are looking for extreme values.
These values are unlikely to be the result of imputation and are also unlikely to be randomly removed from the dataset.

\item We do not compute error for selection queries (\ref{q1}, \ref{q9}, \ref{q10}, \ref{q12}), because the output of the query on clean data is not directly comparable to the output on dirty data.

  \item Imputation performs poorly on the join queries on synthetic data (\ref{q11}, \ref{q13}) because the data is synthetic, so there are no correlations for imputation to find. We included these queries because we wanted to test the performance of our algorithm on queries that include joins, and we were pleased to find that simple joins cause no perceptible performance problems.
\end{itemize}

Finally, we performed CE-CART imputation on the entirety of the ACS table, in order to compare our system to the traditional, pre-processing approach to imputation.
We were unable to complete the full imputation due to resource constraints, but the process ran for more than one hour without terminating.
This time compares favorably to our maximum of 72 seconds for a query with imputation on the ACS data.
Given the varying needs of database users, \ProjectName{} clearly allows flexibility without a significant performance hit.
}

>>>>

%\begin{table*}
%  \centerfloat
%  \input{runtimes}
%    \caption{Base error, percent change in error and and running time for queries
%    with different imputation levels. Base error is the root-mean-square error (RMSE) between the query run on clean
%    data and the query run on dirty data without imputation. Change in error is relative to the base error.}
%  \label{fig:experiments}
%\end{table*}

%\todobox{actually run experiments and rewrite this section}{
All experiments were run on a Macbook Pro with 2 2.9 GHz Intel core i5 processors, 16 GB of main memory,  running OS X El Capitan.

Figure~\ref{fig:runtimes} provides a summary of the performance results. Each query is run in two configurations: quality-optimized and runtime-optimized.
To provide comparison, we also include a table with the average running time required to perform imputation at each of the base tables used in our analysis.
Note that multiple queries perform joins between two or more base tables. In all cases, running a query through \ProjectName{} is cheaper than 
imputing a single base table, with the added flexibility of imputation tailored to the query.

\begin{figure}
\begin{subfigure}{\linewidth}
\includegraphics[width=\columnwidth]{figures/running_times_imputedb.png}
\caption{Query runtimes with \ProjectName{}}
\label{subfig:project-runtime-queries}
\end{subfigure}
~
\begin{subfigure}{\linewidth}
\includegraphics[width=\columnwidth]{figures/running_times_base_tables.png}
\caption{Query runtimes with imputation on base tables}
\label{subfig:base-runtime-queries}
\end{subfigure}
\caption{Comparing \ProjectName{} runtimes for each query (\ref{subfig:project-runtime-queries}) with the runtime for a query when base-table imputations are used (\ref{subfig:base-runtime-queries})}
\end{figure}

Figure~\ref{fig:plantimes} provides a summary of the planning times for each of the queries.  We exclude the planning time for queries that impute at base table, as that
requires no planning. On average, planning constituted 38 percent of total runtime for plans generated by \ProjectName{}.

\begin{figure}
\includegraphics[width=\columnwidth]{figures/planning_times_imputedb.png}
\caption{Planning times for each query}
\label{fig:plantimes}
\end{figure}

Table~\ref{table:rmse} shows the Root-Mean-Squared-Error (RMSE) for \ProjectName{}'s query results when compared to running imputation on the base tables and executing the query on the
cleaned data. Each query was run 10 iterations in both settings and then results from the two approaches were paired up and compared tuple-wise. We average the
squared error across a single query before applying square root. We can see that optimizing for quality indeed reduces the RMSE of query results. In general, the RMSE relative to the
base-imputation approach are acceptable.

\begin{table}
\input{perf_summary.tex}
\caption{Root-Mean-Squared-Error for queries run under different $\alpha$ parameterizations relative to results when imputing on base table}
\label{table:rmse}
\end{table}
